Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2020-10-05 GWLDCI.tex
most of the comments from report Y

The second paper \cite{generalcombinatoricsII} in this series identifies precisely which positroid cell is associated to a given Wilson loop diagram, by giving an algorithm for passing from a Wilson loop diagram to the Grassmann necklace of the associated positroid cell. In this manner we give a direct path from Wilson loop diagram to positroid cell, circumventing the painstaking process of examining sets of bases. With this identification in hand, we also prove that any Wilson loop diagram with $k$ propagators corresponds to a $3k$ dimensional positroid cell, and characterize the volume forms that Wilson loop diagrams associate to each positroid cell. The results of the first paper are not prerequisite for the second paper in this series.

@@

Meanwhile, there is a separate body of work studying SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv, Britto:2005fq, Cachazo:2004kj}. These integrals are calculated via Wilson loops. As noted above, Agarwala and Marin-Amat uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that are realized as an element of the positive Grassmannian $\Gr(k,n)$ \cite{wilsonloop}. In separate papers with Fryer and with Marcott, this work was extended to study these Feynman integrals in terms of positroid cells and the positive Grassmannians \cite{casestudy, non-orientable}. Others have tried different approaches to define a geometric space associated to these integrals in a manner similar to the Amplituhedron \cite{Amplituhedronsquared, HeslopStewart}.

@@

However, the two approaches are not identical: for example, in the case of Wilson loop diagrams with $k=2$ and $n=6$, past work of Agarwala and Fryer shows that the manifold is not contractible \cite{casestudy}. More generally, it has been shown that the spaces defined by the Amplituhedron and by Wilson loop diagrams are not the same \cite{Amplituhedronsquared}. In fact, it is in some ways more natural to consider Deodhar cells of $\mathbb{G}(k, n+1)$ for Wilson loop diagrams; in this case, one gets that the manifold represented by the diagrams is not orientable \cite{non-orientable}.

In this paper we focus on the correspondence between Wilson loop diagrams and positroids, which was first defined in \cite{wilsonloop}. Our main results are as follows. We first show that two Wilson loop diagrams define the same positroid if and only if they only differ by exact subdiagrams (Theorem~\ref{same matroid iff equiv}); equivalently, this says that the map from Wilson loop diagrams to positroid cells is one-to-one exactly on the subset of diagrams with no non-trivial exact subdiagrams. We count the total number of Wilson loop diagrams associated to a positroid cell (Corollary~\ref{number of equiv diagrams}). Finally, we show that enumerating the number of distinct positroid cells associated to Wilson loop diagrams is equivalent to enumerating the number of directions of an associahedron (Theorem~\ref{thm:count inequiv diagrams}).

@@

A Wilson loop diagram can also be represented visually, which provides useful intuition for many of the results that follow. We depict the diagram $(\cP,V)$ as a circle with marked points, called {\em vertices}; these vertices are labeled by $V$ (preserving the cyclic ordering). The arcs between consecutive vertices are called {\em edges}. There are also $k$ wavy lines in the interior of the diagram, each connecting a pair of edges; these depict the propagators. Specifically, a propagator $p =(i,j)$ has one endpoint on the edge joining vertices $i$ and $i+1$ and another endpoint on the edge joining $j$ and $j+1$, where $i+1$ and $j+1$ denote the successor of $i$ and $j$ respectively in the cyclic order on $V$.

@@

\vspace{3pt}
\begin{eg}
The Wilson loop diagram $W$ in Example~\ref{eg of WLD} is an example of an admissible diagram. It has a single non-supporting vertex, namely the vertex labeled 7, so we have $F(\emptyset) = \{7\}$. For the propagator $p = (5, 8)$, we have $V(p) = \{5,6,8,1\}$ and $F(p) = \{6, 7,8\}$: the vertices 1 and 5 are excluded from $F(p)$ since they support other propagators as well, and 7 is included (despite not being in the support of $p$) since it supports no propagators at all. Finally, the subdiagram $(\{(2, 4), (1, 4)\}, \{1, 2, 3, 4, 5\})$ is an example of an exact subdiagram; in fact it is the only non-trivial exact subdiagram in $W$.
\end{eg}

@@

The matrix $C(W)$ allows us to associate a matroid to each Wilson loop diagram. We first give a quick summary of the matroid terminology that we will need; it is not intended as a comprehensive introduction to matroids and the interested reader is referred to \cite{OxleyMatroidBook}

A {\em matroid} $M = (E,\cB)$ consists of a finite ground set $E$ and a non-empty family $\cB$ of subsets of $E$, such that elements of $\cB$ satisfy the {\em basis exchange property}: for any distinct $B_1,B_2 \in \cB$ and any $a \in B_1 \setminus B_2$, there exists some $b \in B_2 \setminus B_1$ such that $(B_1 \setminus \{a\})\cup \{b\} \in \cB$ as well. The elements of $\cB$ are called the {\em bases} of the matroid. Note that the basis exchange property immediately implies that all bases have the same size.

@@

A {\em circuit} in $M$ is a minimally dependent set. That is, it is a set $C \subseteq E$ such that $C$ is dependent but $C \setminus \{e\}$ is independent for any $e \in C$. A union of circuits is called a {\em cycle}. On the other hand, a {\em flat} is a maximally dependent set, i.e. a set $F \subseteq E$ such that $\rk(F \cup \{e\}) = \rk(F) + 1$ for any $e \in E \setminus F$. Unsurprisingly, a {\em cyclic flat} is a set which is both a flat and a cycle. The set of circuits in a matroid uniquely defines that matroid, as does the set of flats; thus one could specify a matroid by listing either its independent sets, its bases, its circuits, or its flats.

@@

\begin{lem}\label{lem facts about WLD matroids}
Let $W = (\cP,[n])$ be a weakly admissible Wilson loop diagram. Then:
\begin{enumerate}
\item The rank of a set $V \subseteq [n]$ is bounded above by $\min\{|V|,|\Prop(V)|\}$, with $\rk(V) = |V|$ if and only if $V$ is an independent set.
%\item Let $v \in [n]$ and $q \in \Prop(v)$. Then for any $V \subset [n]$ such that $q \not \in \Prop(V)$, we have $\rk(V \cup \{v\}) = \rk(V) +1$, i.e. adding a vertex that supports a new propagator to a set increases the rank of the set.
\item If $C \subseteq [n]$ is a cycle in $M(W)$, then $\rk(C) = |Prop(C)|$.
\item If $[n]$ can be partitioned into at least two non-empty sets, each of which support disjoint sets of propagators and in such a way that we obtain a partition of the propagator set, i.e. if we have \bas [n] = \sqcup_i V(P_i) \ \textrm{ such that } \ \sqcup P_i = \cP; \   \text{ with }V(P_i) \cap V(P_j) = \emptyset \text{ and } P_i \cap P_j = \emptyset \quad \forall i \neq j\;, \eas then the matroid $M(W)$ is separable, i.e. \bas M(W) = \bigoplus_i M(P_i, V(P_i)) \;.\eas 
\item The set $F(P)$ from Definition~\ref{VPropdfn} is a flat of $M( W)$, which we call the \emph{propagator flat} of $P$.
\item If $F$ is a cyclic flat of $M(W)$, then it is a propagator flat.
\end{enumerate}
\end{lem}
\begin{proof}
The first part of (1) is \cite[Equation (9)]{wilsonloop} and surrounding discussion, and the second part is standard matroid theory. (2) is \cite[Lemma 3.27]{wilsonloop}. (3) is a direct consequence of \cite[Lemma 3.20]{wilsonloop} and the fact that $F(P_1)^c = V(P_1^c)$, and (5) is given in \cite[Lemma 3.28]{wilsonloop}.

To prove (4), we need to show that $F(P)$ is maximally dependent. If $F(P) = [n]$ then this is automatic, so suppose not and let $v \in [n] \setminus F(P)$. In other words, $v \in V(P^c)$ and so $v$ supports some propagator $q \not\in P$.  Let $S \subseteq F(P)$ be an independent set of maximal size. Then $\Prop(S) \subseteq P$ (from the definition of $F(P)$) and no subset of $S$ supports fewer propagators than the number of vertices it contains (by Theorem~\ref{thm WLD defines matroid}). Since $v$ supports a new propagator $q \not\in P$, the set $S \cup \{v\} \subseteq F(P) \cup \{v\}$ also satisfies this independence condition.

@@

\begin{rmk}
By point (4) of Lemma~\ref{lem facts about WLD matroids}, the set $F(\emptyset)$ of non-supporting vertices in $W$ is a flat; indeed, it is the unique flat of rank 0 in $M(W)$. Every element of $F(\emptyset)$ defines a circuit of size 1, and all other circuits of $M(W)$ must contain at least two elements.
\end{rmk}

@@

In this section we develop this analysis further, showing the converse of this result from loc. cit. (see Theorem \ref{same matroid iff equiv}). We also enumerate the number of diagrams that map to a particular matroid in Corollary \ref{number of equiv diagrams}. To do this, we establish a connection between Wilson loop diagrams and polygon dissections in subsection \ref{sec: polygon partitions}, and uncover some matroidal properties of exact subdiagrams in subsection \ref{sec: exact diagram matroidal props}. This includes the fact that exact subdiagrams correspond to uniform matroids (Theorem \ref{exactuniformthm}), and that the associated matroid of an exact subdiagram can be written as a contraction by the complementary propagator flat (Theorem \ref{exact diagrams contractions}).

@@

\begin{dfn}\label{WLDtriangulationdfn}
  Let $W = (\cP, [n])$ be a weakly admissible Wilson loop diagram.  The \emph{polygon dissection} associated to $W$, denoted $\tau(W)$, is defined as follows.
  \begin{itemize}
  \item The vertices of $\tau(W)$ correspond to the edges of $W$.
  \item Labeling the vertices of $\tau(W)$ with the edge number of $W$ gives a cyclic order to the vertices. Connecting consecutive vertices gives a graph theoretic cycle called the polygon of $\tau(W)$.
  \item Each propagator of $W$ defines a diagonal edge of $\tau(W)$; specifically,  a propagator $(i,j) \in \cP$ defines a diagonal connecting the vertices $i$ and $j$ in $\tau(W)$.
  \end{itemize}
\end{dfn}

@@

\begin{dfn}
  Let $W$ be a weakly admissible Wilson loop diagram and $\tau(W)$ its polygon dissection. A \emph{triangulated piece} of $\tau(W)$ is a 2-connected subgraph of $\tau(W)$ which is a triangulation. We will take the convention that a subgraph consisting of a single diagonal edge is called a \emph{trivial} triangulated piece.
A {\em maximal} triangulated piece is one which is not contained in any strictly larger triangulated piece.
\end{dfn}

@@

\begin{dfn}
 A {\em decomposition} of a polygon dissection $\tau(W)$ is a set of 2-connected induced subgraphs of $\tau(W)$ which partition the edges of $\tau(W)$.  
\end{dfn}

@@

\begin{proof}
Let us first record a few standard facts about polygon triangulations (that is, about triangulations with all vertices on the outer face).  If such a triangulation has $n$ vertices then it has $n$ edges on the polygon (that is, on the outer face) and $n-3$ edges which are not.  No planar simple graph with the same vertices and the same outer face can have more edges than a triangulation, and every such simple graph with $n-3$ edges off the outer face is a triangulation.

Since $W$ is weakly admissible, by Lemma \ref{tausimpleplanarlem} $\tau(W)$ is a simple graph. Let $t$ be a triangulated piece of the decomposition of $\tau(W)$ given in Lemma \ref{decompositionlem}; note that $t$ cannot be equal to $\tau(W)$ by the definition of admissible diagrams.

@@

Each of these $m-3+j$ edges corresponds to a propagator in $W$; call this set of propagators $P$.  Next we count the size of $V(P)$.  Each of the $m$ vertices in the outer face of $t$ corresponds to an edge of $W$. These $m$ edges define $j$ connected components of the outer polygon of $W$. Thus the set $V(P)$ has $m+j$ vertices.  In other words, 
\[|V(P)| = m+j = |P| +3\;.\]
Thus the subdiagram $(P,V(P))$ defined by $t$ is exact.

@@

Now suppose there exists some $w \in  F(\Prop(C)) \setminus C$. Let $B$ be an independent subset of $C$ of maximal rank; we first show that $B \cup \{w\}$ is a dependent set in $M(W)$. By Lemma \ref{lem facts about WLD matroids}, we have
\begin{equation}\label{eq:equality in lem decompose flat} |\Prop(C)| = \rk(C) = |B| = \rk(B) .\end{equation}
Note that $\rk(B) \leq |\Prop(B)|$, and $B \subseteq C$ implies that $\Prop(B) \subseteq \Prop(C)$; combining these facts with equation \eqref{eq:equality in lem decompose flat}, it follows that $\Prop(B) = \Prop(C)$. Combining this with the fact that $w \in F(\Prop(C))$, we have $\Prop(B\cup \{w\}) \subseteq \Prop(B)$. We can now say something about $\rk(B\cup \{w\})$: by Lemma \ref{lem facts about WLD matroids}, we have 
\[
\rk (B \cup \{w\}) \leq \min\{|B\cup \{w\}|, |\Prop(B\cup \{w\})|\} \leq \min\{|B|+1,|\Prop(B)|\},
\]
which becomes 
\[\rk(B\cup \{w\}) \leq \min\{|B|+1,|B|\} = |B| < |B \cup\{w\}|\]
via equation \eqref{eq:equality in lem decompose flat}. Therefore $B\cup\{w\}$ is a dependent set, which implies that $B\cup \{w\}$ contains a circuit $D$. In particular we must have $w \in D$ (since $B$ was an independent set), and since we have already shown that $C \subseteq F(\Prop(C))$, we also have the chain of inclusions 
\[D \subseteq B\cup \{w\} \subseteq F(\Prop(C)) \subseteq F.\]
Since $C$ was the union of all circuits in $F$, we finally obtain $w \in C$, which is a contradiction. This completes the proof of (1).

@@

Now suppose that $F(\emptyset) = \emptyset$; we must show that $F\setminus C$ is a flat. For any $e \not\in F$ we certainly have $\rk((F\setminus C)\cup\{e\}) = \rk(F\setminus C) +1$, since $F$ is a flat. Now let $e \in C$, and suppose that $\rk((F\setminus C)\cup\{e\}) = \rk(F\setminus C)$. This implies that $(F\setminus C)\cup \{e\}$ is dependent, and hence contains a circuit. Since $F(\emptyset) = \emptyset$ this circuit must contain at least two elements: one of these elements is $e$, but any others must come from $F \setminus C$. This contradicts the fact that $C$ was the union of all circuits in $F$.

@@

We know from Lemma~\ref{lem facts about WLD matroids} that $F(P^c)$ is always a flat, so for part (1) we only need to show that $F(P^c)$ is dependent.

@@

For part (2), suppose that $(P, V(P))$ is a maximal exact subdiagram of $W$. By Lemma \ref{lem decompose flat}(1) and Corollary \ref{classifyflats} we can decompose $F(P^c)$ as
\begin{equation}\label{eq decompose F}V(P)^c = F(P^c) = C \sqcup S = F(\Prop(C)) \sqcup S,\end{equation}
where $C$ is the largest cyclic flat contained in $F(P^c)$ and $S$ is an independent set. We show that $\Prop(C) = P^c$, thus forcing $S$ to be empty and implying that $F(P^c) = C$, i.e. $F(P^c)$ is a cyclic flat. Note that $C$ must be non-empty since $F(P^c)$ is non-empty and dependent (by part (1)).

By similar arguments to Lemma~\ref{lem decompose flat}, the fact that $S$ is independent and that $C$ is the union of all circuits in $F(P^c)$ implies that every element of $S$ is independent of $C$. Further, we know that $\rk(C) = |\Prop(C)|$ from Lemma \ref{lem facts about WLD matroids}(2), and $\rk(S) = |S|$ from Lemma \ref{lem facts about WLD matroids}(1) and the independence of $S$.  Combining these gives
\ba \rk\big(F(P^c)\big) = \rk(C) + \rk(S)  = |\Prop(C)| + |S| \leq |\Prop(F(P^c))|\;,\label{rank of comp flat} \ea 
where the inequality follows from the bound on ranks of vertex sets in Lemma \ref{lem facts about WLD matroids}(1). For ease of future notation, write $Q = \Prop(C)^c$.

@@

It remains to show that $S$ is an independent set in the subdiagram $(P, V(P))$, so for a contradiction suppose otherwise. Then by Theorem \ref{thm WLD defines matroid} there is a subset $U \subseteq S$ such that $|\Prop(U)| < |U|$. Then the remaining propagators $P \setminus \Prop(U)$ are supported entirely on $V(P) \setminus U$, i.e. \bas  V(P \setminus \Prop(U)) \subseteq V(P) \setminus U\; .\eas Comparing sizes of these sets, we get \bas |V(P \setminus \Prop(U))| \leq |V(P)| - |U| = |P| +3 -|U| < |P| +3 -|\Prop(U)| = |P \setminus \Prop(U)| +3\; ,\eas which violates admissibility. Therefore $S$ must be independent in $(P, V(P))$ after all, and (as described above) it follows that $\rk(F(R^c)) = |R^c|$.

@@

\begin{proof}
It follows directly from the definitions that a matroid of rank $r$ is uniform if and only if all circuits have rank $r$; we therefore focus on the circuits of $M(W')$.

@@

By Theorem \ref{exactuniformthm}, $M(R, V(R))$ and $M(R',V(R'))$ are both uniform matroids of rank $|R| = |R'|$; in particular, this means that any subset of $V(R)$ is independent in both $M(W)$ and $M(W')$.

@@

By Lemma~\ref{maxexactcomplementrank} we know that $\rk(F) = |P^c|$; however, by Lemma~\ref{lem facts about WLD matroids} we also know that the rank of a cycle is equal to the number of propagators supported on that cycle, so $\rk(F) = |\Prop_{W'}(F)| = |Q^c|$ as well. Since $(P,V(P))$ is exact and we have shown that $V(P) = V(Q)$ and $|P| = |Q|$, it follows that $|V(Q)| = |Q| + 3$, i.e.  $(Q,V(Q))$ is exact in $W'$.

@@

As with many of the interesting and well-studied polytopes in combinatorics, the associahedron can be characterized in many different ways; here is one definition.

>>>>

The second paper \cite{generalcombinatoricsII} in this series identifies precisely which positroid cell is associated to a given Wilson loop diagram, by giving an algorithm for passing from a Wilson loop diagram to the Grassmann necklace of the associated positroid cell. In this manner we give a direct path from a Wilson loop diagram to a positroid cell, circumventing the painstaking process of examining sets of bases. With this identification in hand, we also prove that any Wilson loop diagram with $k$ propagators corresponds to a $3k$ dimensional positroid cell, and characterize the volume forms that Wilson loop diagrams associate to each positroid cell. The results of the first paper are not prerequisite for the second paper in this series.

@@

Meanwhile, there is a separate body of work studying SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv, Britto:2005fq, Cachazo:2004kj}. These integrals are calculated via Wilson loops. As noted above, Agarwala and Marin-Amat uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that are realized as an element of the positive Grassmannian $\Gr(k,n)$ \cite{wilsonloop}. In separate papers with Fryer and with Marcott, this work was extended to study these Feynman integrals in terms of positroid cells and the positive Grassmannian \cite{casestudy, non-orientable}. Others have tried different approaches to define a geometric space associated to these integrals in a manner similar to the Amplituhedron \cite{Amplituhedronsquared, HeslopStewart}.

@@

However, the two approaches are not identical: for example, in the case of Wilson loop diagrams with $k=2$ and $n=6$, past work of Agarwala and Fryer shows that the manifold is not contractible \cite{casestudy}. More generally, it has been shown that the spaces defined by the Amplituhedron and by Wilson loop diagrams are not the same \cite{Amplituhedronsquared}. In fact, it is in some ways more natural to consider Deodhar cells of $\mathbb{G}(k, n+1)$ for Wilson loop diagrams.  In this case, one gets that the manifold represented by the diagrams is not orientable \cite{non-orientable}.

In this paper we focus on the correspondence between Wilson loop diagrams and positroids, which was first defined in \cite{wilsonloop}. Our main results are as follows. We first show that two Wilson loop diagrams define the same positroid if and only if they only differ by exact subdiagrams (Theorem~\ref{same matroid iff equiv}). Equivalently, this says that the map from Wilson loop diagrams to positroid cells is one-to-one exactly on the subset of diagrams with no non-trivial exact subdiagrams. We count the total number of Wilson loop diagrams associated to a positroid cell (Corollary~\ref{number of equiv diagrams}). Finally, we show that enumerating the number of distinct positroid cells associated to Wilson loop diagrams is equivalent to enumerating the number of directions of an associahedron (Theorem~\ref{thm:count inequiv diagrams}).

@@

A Wilson loop diagram can also be represented visually, which provides useful intuition for many of the results that follow. We depict the diagram $(\cP,V)$ as a circle with marked points, called {\em vertices}.  These vertices are labeled by $V$ (preserving the cyclic ordering). The arcs between consecutive vertices are called {\em edges}. There are also $k$ wavy lines in the interior of the diagram, each connecting a pair of edges.  These depict the propagators. Specifically, a propagator $p =(i,j)$ has one endpoint on the edge joining vertices $i$ and $i+1$ and another endpoint on the edge joining $j$ and $j+1$, where $i+1$ and $j+1$ denote the successor of $i$ and $j$ respectively in the cyclic order on $V$.

@@

\vspace{3pt}
\begin{eg}
The Wilson loop diagram $W$ in Example~\ref{eg of WLD} is an example of an admissible diagram. It has a single non-supporting vertex, namely the vertex labeled 7, so we have $F(\emptyset) = \{7\}$. For the propagator $p = (5, 8)$, we have $V(p) = \{5,6,8,1\}$ and $F(p) = \{6, 7,8\}$: the vertices 1 and 5 are excluded from $F(p)$ since they support other propagators as well, and 7 is included (despite not being in the support of $p$) since it supports no propagators at all. Finally, the subdiagram $(\{(2, 4), (1, 4)\}, \{1, 2, 3, 4, 5\})$ is an example of an exact subdiagram.  In fact it is the only non-trivial exact subdiagram in $W$.
\end{eg}

@@

The matrix $C(W)$ allows us to associate a matroid to each Wilson loop diagram. We first give a quick summary of the matroid terminology that we will need.  It is not intended as a comprehensive introduction to matroids and the interested reader is referred to \cite{OxleyMatroidBook}.

A {\em matroid} $M = (E,\cB)$ consists of a finite ground set $E$ and a non-empty family $\cB$ of subsets of $E$, such that elements of $\cB$ satisfy the {\em basis exchange property}: for any distinct $B_1,B_2 \in \cB$ and any $a \in B_1 \setminus B_2$, there exists some $b \in B_2 \setminus B_1$ such that $(B_1 \setminus \{a\})\cup \{b\} \in \cB$. The elements of $\cB$ are called the {\em bases} of the matroid. Note that the basis exchange property immediately implies that all bases have the same size.

@@

A {\em circuit} in $M$ is a minimally dependent set. That is, it is a set $C \subseteq E$ such that $C$ is dependent but $C \setminus \{e\}$ is independent for any $e \in C$. A union of circuits is called a {\em cycle}. On the other hand, a {\em flat} is a maximally dependent set, i.e. a set $F \subseteq E$ such that $\rk(F \cup \{e\}) = \rk(F) + 1$ for any $e \in E \setminus F$. Unsurprisingly, a {\em cyclic flat} is a set which is both a flat and a cycle. The set of circuits in a matroid uniquely defines that matroid, as does the set of flats.  Thus one could specify a matroid by listing either its independent sets, its bases, its circuits, or its flats.

@@

\begin{lem}\label{lem facts about WLD matroids}
Let $W = (\cP,[n])$ be a weakly admissible Wilson loop diagram. Then:
\begin{enumerate}
\item The rank of a set $V \subseteq [n]$ is bounded above by $\min\{|V|,|\Prop(V)|\}$, with $\rk(V) = |V|$ if and only if $V$ is an independent set.
%\item Let $v \in [n]$ and $q \in \Prop(v)$. Then for any $V \subset [n]$ such that $q \not \in \Prop(V)$, we have $\rk(V \cup \{v\}) = \rk(V) +1$, i.e. adding a vertex that supports a new propagator to a set increases the rank of the set.
\item If $C \subseteq [n]$ is a cycle in $M(W)$, then $\rk(C) = |Prop(C)|$.
\item If $[n]$ can be partitioned into at least two non-empty sets, each of which support disjoint sets of propagators and in such a way that we obtain a partition of the propagator set, i.e. if we have \bas [n] = \sqcup_i V(P_i) \ \textrm{ such that } \ \sqcup P_i = \cP; \   \text{ with }V(P_i) \cap V(P_j) = \emptyset \text{ and } P_i \cap P_j = \emptyset \quad \forall i \neq j\;, \eas then the matroid $M(W)$ is separable, i.e. \bas M(W) = \bigoplus_i M(P_i, V(P_i)) \;.\eas 
\item The set $F(P)$ from Definition~\ref{VPropdfn} is a flat of $M( W)$, which we call the \emph{propagator flat} of $P$.
\item If $F$ is a cyclic flat of $M(W)$, then it is a propagator flat.
\end{enumerate}
\end{lem}
\begin{proof}
The first part of item 1 is \cite[Equation (9)]{wilsonloop} and surrounding discussion, and the second part is standard matroid theory. Item 2 is \cite[Lemma 3.27]{wilsonloop}. Item 3 is a direct consequence of \cite[Lemma 3.20]{wilsonloop} and the fact that $F(P_1)^c = V(P_1^c)$, and item 5 is given in \cite[Lemma 3.28]{wilsonloop}.

To prove item 4, we need to show that $F(P)$ is maximally dependent. If $F(P) = [n]$ then this is automatic, so suppose not and let $v \in [n] \setminus F(P)$. In other words, $v \in V(P^c)$ and so $v$ supports some propagator $q \not\in P$.  Let $S \subseteq F(P)$ be an independent set of maximal size. Then $\Prop(S) \subseteq P$ (from the definition of $F(P)$) and no subset of $S$ supports fewer propagators than the number of vertices it contains (by Theorem~\ref{thm WLD defines matroid}). Since $v$ supports a new propagator $q \not\in P$, the set $S \cup \{v\} \subseteq F(P) \cup \{v\}$ also satisfies this independence condition.

@@

\begin{rmk}
By item 4 of Lemma~\ref{lem facts about WLD matroids}, the set $F(\emptyset)$ of non-supporting vertices in $W$ is a flat. Indeed, it is the unique flat of rank 0 in $M(W)$. Every element of $F(\emptyset)$ defines a circuit of size 1, and all other circuits of $M(W)$ must contain at least two elements.
\end{rmk}

@@

In this section we develop this analysis further, showing the converse of this result from \cite{wilsonloop} (see Theorem \ref{same matroid iff equiv}). We also enumerate the number of diagrams that map to a particular matroid in Corollary \ref{number of equiv diagrams}. To do this, we establish a connection between Wilson loop diagrams and polygon dissections in Subsection \ref{sec: polygon partitions}, and uncover some matroidal properties of exact subdiagrams in Subsection \ref{sec: exact diagram matroidal props}. This includes the fact that exact subdiagrams correspond to uniform matroids (Theorem \ref{exactuniformthm}), and that the associated matroid of an exact subdiagram can be written as a contraction by the complementary propagator flat (Theorem \ref{exact diagrams contractions}).

@@

\begin{dfn}\label{WLDtriangulationdfn}
  Let $W = (\cP, [n])$ be a weakly admissible Wilson loop diagram.  The \emph{polygon dissection} associated to $W$, denoted $\tau(W)$, is defined as follows.
  \begin{itemize}
  \item The vertices of $\tau(W)$ correspond to the edges of $W$.
  \item Labeling the vertices of $\tau(W)$ with the edge number of $W$ gives a cyclic order to the vertices. Connecting consecutive vertices gives a graph theoretic cycle called the polygon of $\tau(W)$.
  \item Each propagator of $W$ defines a diagonal edge of $\tau(W)$.  Specifically,  a propagator $(i,j) \in \cP$ defines a diagonal connecting the vertices $i$ and $j$ in $\tau(W)$.
  \end{itemize}
\end{dfn}

@@

\begin{dfn}
  Let $W$ be a weakly admissible Wilson loop diagram and $\tau(W)$ its polygon dissection. A \emph{triangulated piece} of $\tau(W)$ is a connected, bridgeless subgraph of $\tau(W)$ which is a triangulation. We will take the convention that a subgraph consisting of a single diagonal edge is called a \emph{trivial} triangulated piece.
A {\em maximal} triangulated piece is one which is not contained in any strictly larger triangulated piece.
\end{dfn}

@@

\begin{dfn}
 A {\em decomposition} of a polygon dissection $\tau(W)$ is a set of bridgeless, connected, induced subgraphs of $\tau(W)$ which partition the edges of $\tau(W)$.  
\end{dfn}

@@

\begin{proof}
Let us first record a few standard facts about polygon triangulations (that is, about triangulations with all vertices on the outer face).  If such a triangulation has $n$ vertices then it has $n$ edges on the polygon (that is, on the outer face) and $n-3$ edges which are not.  No planar simple graph with the same vertices and the same outer face can have more edges than a triangulation, and every such simple graph with $n-3$ edges off the outer face is a triangulation.\todo{Referee Y wants a ref for these standard things.}

Since $W$ is weakly admissible, by Lemma \ref{tausimpleplanarlem} $\tau(W)$ is a simple graph. Let $t$ be a triangulated piece of the decomposition of $\tau(W)$ given in Lemma \ref{decompositionlem}.  Note that $t$ cannot be equal to $\tau(W)$ by the definition of admissible diagrams.

@@

Each of these $m-3+j$ edges corresponds to a propagator in $W$.  Call this set of propagators $P$.  Next we count the size of $V(P)$.  Each of the $m$ vertices in the outer face of $t$ corresponds to an edge of $W$. These $m$ edges define $j$ connected components of the outer polygon of $W$. Thus the set $V(P)$ has $m+j$ vertices.  In other words, 
\[|V(P)| = m+j = |P| +3\;.\]
Thus the subdiagram $(P,V(P))$ defined by $t$ is exact.

@@

Now suppose there exists some $w \in  F(\Prop(C)) \setminus C$. Let $B$ be an independent subset of $C$ of maximal rank. We first show that $B \cup \{w\}$ is a dependent set in $M(W)$. By Lemma \ref{lem facts about WLD matroids}, we have
\begin{equation}\label{eq:equality in lem decompose flat} |\Prop(C)| = \rk(C) = |B| = \rk(B) .\end{equation}
Note that $\rk(B) \leq |\Prop(B)|$, and $B \subseteq C$ implies that $\Prop(B) \subseteq \Prop(C)$. Combining these facts with equation \eqref{eq:equality in lem decompose flat}, it follows that $\Prop(B) = \Prop(C)$. Combining this with the fact that $w \in F(\Prop(C))$, we have $\Prop(B\cup \{w\}) \subseteq \Prop(B)$. We can now say something about $\rk(B\cup \{w\})$: by Lemma \ref{lem facts about WLD matroids}, we have 
\[
\rk (B \cup \{w\}) \leq \min\{|B\cup \{w\}|, |\Prop(B\cup \{w\})|\} \leq \min\{|B|+1,|\Prop(B)|\},
\]
which becomes 
\[\rk(B\cup \{w\}) \leq \min\{|B|+1,|B|\} = |B| < |B \cup\{w\}|\]
via equation \eqref{eq:equality in lem decompose flat}. Therefore $B\cup\{w\}$ is a dependent set, which implies that $B\cup \{w\}$ contains a circuit $D$. In particular we must have $w \in D$ (since $B$ was an independent set), and since we have already shown that $C \subseteq F(\Prop(C))$, we also have the chain of inclusions 
\[D \subseteq B\cup \{w\} \subseteq F(\Prop(C)) \subseteq F.\]
Since $C$ was the union of all circuits in $F$, we finally obtain $w \in C$, which is a contradiction. This completes the proof of (1).

@@

Now suppose that $F(\emptyset) = \emptyset$. We must show that $F\setminus C$ is a flat. For any $e \not\in F$ we certainly have $\rk((F\setminus C)\cup\{e\}) = \rk(F\setminus C) +1$, since $F$ is a flat. Now let $e \in C$, and suppose that $\rk((F\setminus C)\cup\{e\}) = \rk(F\setminus C)$. This implies that $(F\setminus C)\cup \{e\}$ is dependent, and hence contains a circuit. Since $F(\emptyset) = \emptyset$ this circuit must contain at least two elements: one of these elements is $e$, but any others must come from $F \setminus C$. This contradicts the fact that $C$ was the union of all circuits in $F$.

@@

We know from Lemma~\ref{lem facts about WLD matroids} that $F(P^c)$ is always a flat, so for item 1 we only need to show that $F(P^c)$ is dependent.

@@

For item 2, suppose that $(P, V(P))$ is a maximal exact subdiagram of $W$. By Lemma \ref{lem decompose flat} item 1 and Corollary \ref{classifyflats} we can decompose $F(P^c)$ as
\begin{equation}\label{eq decompose F}V(P)^c = F(P^c) = C \sqcup S = F(\Prop(C)) \sqcup S,\end{equation}
where $C$ is the largest cyclic flat contained in $F(P^c)$ and $S$ is an independent set. We show that $\Prop(C) = P^c$, thus forcing $S$ to be empty and implying that $F(P^c) = C$, i.e. $F(P^c)$ is a cyclic flat. Note that $C$ must be non-empty since $F(P^c)$ is non-empty and dependent (by item 1).

By similar arguments to Lemma~\ref{lem decompose flat}, the fact that $S$ is independent and that $C$ is the union of all circuits in $F(P^c)$ implies that every element of $S$ is independent of $C$. Further, we know that $\rk(C) = |\Prop(C)|$ from Lemma \ref{lem facts about WLD matroids}(2), and $\rk(S) = |S|$ from Lemma \ref{lem facts about WLD matroids} item 1 and the independence of $S$.  Combining these gives
\ba \rk\big(F(P^c)\big) = \rk(C) + \rk(S)  = |\Prop(C)| + |S| \leq |\Prop(F(P^c))|\;,\label{rank of comp flat} \ea 
where the inequality follows from the bound on ranks of vertex sets in Lemma \ref{lem facts about WLD matroids} item 1. For ease of future notation, write $Q = \Prop(C)^c$.

@@

It remains to show that $S$ is an independent set in the subdiagram $(P, V(P))$.  By way of contradiction suppose otherwise. Then by Theorem \ref{thm WLD defines matroid} there is a subset $U \subseteq S$ such that $|\Prop(U)| < |U|$. Then the remaining propagators $P \setminus \Prop(U)$ are supported entirely on $V(P) \setminus U$, i.e. \bas  V(P \setminus \Prop(U)) \subseteq V(P) \setminus U\; .\eas Comparing sizes of these sets, we get \bas |V(P \setminus \Prop(U))| \leq |V(P)| - |U| = |P| +3 -|U| < |P| +3 -|\Prop(U)| = |P \setminus \Prop(U)| +3\; ,\eas which violates admissibility. Therefore $S$ must be independent in $(P, V(P))$ after all, and (as described above) it follows that $\rk(F(R^c)) = |R^c|$.

@@

\begin{proof}
It follows directly from the definitions that a matroid of rank $r$ is uniform if and only if all circuits have rank $r$. We therefore focus on the circuits of $M(W')$.

@@

By Theorem \ref{exactuniformthm}, $M(R, V(R))$ and $M(R',V(R'))$ are both uniform matroids of rank $|R| = |R'|$. In particular, this means that any subset of $V(R)$ is independent in both $M(W)$ and $M(W')$.

@@

By Lemma~\ref{maxexactcomplementrank} we know that $\rk(F) = |P^c|$. However, by Lemma~\ref{lem facts about WLD matroids} we also know that the rank of a cycle is equal to the number of propagators supported on that cycle, so $\rk(F) = |\Prop_{W'}(F)| = |Q^c|$ as well. Since $(P,V(P))$ is exact and we have shown that $V(P) = V(Q)$ and $|P| = |Q|$, it follows that $|V(Q)| = |Q| + 3$, i.e.  $(Q,V(Q))$ is exact in $W'$.

@@

As with many of the interesting and well-studied polytopes in combinatorics, the associahedron can be characterized in many different ways. Here is one definition.
